---
title: "New package in CRAN: PkgsFromFiles"
subtitle: ""
author: "Marcelo S. Perlin"
date: '2018-10-13'
image:
  caption: '/post/2018-10-13-NewPackage-PkgsFromFiles_files/figure-html/unnamed-chunk-6-1.png'
  focal_point: ''
categories: [R, PkgsFromFiles]   
tags: [R, PkgsFromFiles]   
---

```{r, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```


Its been a while since I develop a CRAN package and this weekend I decided
to work on a idea I had some time ago. The result is package
`PkgsFromFiles`.

When working with different computers at home or work, one of the
problems I have is installing missing packages across different
computers. As an example, a script that works in my **work computer** may
not work in my **home computer**. This is specially annoying when I have a
fresh install of the operating system or R. In this case, I must
manually install all packages, case by case. Instead of focusing on the
script at hand, I spend considerable time finding and installing missing
packages. When using laptops for teaching R, many times I had to wait
for the installation of a package before continuing the class. With my
new package, PkgsFromFiles, I can scan any folder of my computer and
install all necessary packages **before** using them, as we will soon
learn.

One of the available solutions to this problem is to use package
[pacman](https://CRAN.R-project.org/package=pacman). It includes
function `p_load` that will check if a package is available and, if not,
install it from CRAN. However, for me, I like using `library` and
`require` as it is consistent with my code format. Also, in a fresh R
install, I rather install all my required packages in a single run so
that I don't have to wait later.

Package PkgsFromFiles solves this issue by finding and parsing all R
related files (.R, .Rmd, .Rnw) from a given folder. It finds all calls
to library() and require() and installs all packages that are not
available locally.

# Installation

```{r, eval=FALSE}
# from cran (soon!)
install.packages('PkgsFromFiles')

# from github
if (!require(devtools)) install.packages('devtools')
devtools::install_github('msperlin/PkgsFromFiles')
```

# Usage

The main function of the package is `pff_find_and_install_pkgs`, which will search and install missing packages from R files at a given directory. As an example, we'll use my research folder from Dropbox. It contains all R scripts I have ever used in my research work. Let's try it out:


```{r}
# Evaluation is disable so it passes CRAN CHECKS, but you should be able to run it in your computer
library(PkgsFromFiles)

# target folder
my.dir <- '~/Dropbox/01-Pesquisa/'

df <- pff_find_and_install_pkgs(folder.in = my.dir)

```

As you can see, function `pff_find_and_install_pkgs` will find all R related files recursively in the given folder. In this case, I have all packages locally so no installation was required. A summary in text is shown at the end of execution. 

The output of the function is a dataframe with the details of the operation. Have a look:

```{r}
dplyr::glimpse(df)
```

The package also includes function `pff_find_R_files_from_folder`, which will find all packages used in R related files from a given folder. It outputs a dataframe with several information about packages used in the found scripts. 

```{r}
df.files <- pff_find_R_files_from_folder(folder.in = my.dir)
dplyr::glimpse(df.files)
```

I also wrote a simple function for plotting the most used packages for a given folder:

```{r}
# target folder
my.dir <- '~/Dropbox/01-Pesquisa/'

# plot most used pkgs
p <- pff_plot_summary_pkgs(folder.in = my.dir)
print(p)

```

As you can see, I'm a big fan of the `tidyverse`!

Hope you guys find the package useful! Fell free to send any question to the comment section of the post or my email (marceloperlin@gmail.com).  
