---
title: "Update to GetLattesData"
subtitle: ""
author: "Marcelo S. Perlin"
date: 2018-10-11
categories: ["R", "GetLattesData"]
tags: [R, GetLattesData]
---


```{r, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

Last year I released `GetLattesData`. This package is very handy for anyone that researches bibliometric data of Brazilian scholars. You could easily import the whole academic history of any researcher registered at the platform. More details about Lattes and `GetLattesData` in the this [post](https://msperlin.github.io/2017-09-04-Package-GetLattesData/). 

However, a couple months ago CNPQ introduced a captcha in the webpage. This made it impossible to download the xml files directly, breaking my code. It seems that those changes are now permanent. The update to GetLattesData will address this issue by asking the user to download the files manually and input its location to function `gld_get_lattes_data_from_zip`. Unfortunately, one can no longer download the files automatically. 

Next I provide an example of usage from the vignette:


```{r}
library(GetLattesData)

# get files from pkg (you can download from other researchers in lattes website)
f.in <- c(system.file('extdata/3262699324398819.zip', package = 'GetLattesData'),
          system.file('extdata/8373564643000623.zip', package = 'GetLattesData'))

# set qualis
field.qualis = 'ADMINISTRAÇÃO PÚBLICA E DE EMPRESAS, CIÊNCIAS CONTÁBEIS E TURISMO'

# get data
l.out <- gld_get_lattes_data_from_zip(f.in, 
                                      field.qualis = field.qualis )

```

The output `my.l` is a list with the following dataframes:

```{r}
names(l.out)
```

The first is a dataframe with information about researchers:

```{r}
tpesq <- l.out$tpesq
str(tpesq)
```

The second dataframe contains information about all published publications, including Qualis and SJR:

```{r}
dplyr::glimpse(l.out$tpublic.published)
```

Other dataframes in `l.out` included information about accepted papers, supervisions, books and conferences.

## An application of  `GetLattesData`

`GetLattesData` makes it easy to create academic reports for a large number of researchers. See next, where I plot the number of publications for each researcher, conditioning on Qualis ranking.

```{r}
tpublic.published <- l.out$tpublic.published

library(ggplot2)

p <- ggplot(tpublic.published, aes(x = qualis)) +
  geom_bar(position = 'identity') + facet_wrap(~name) +
  labs(x = paste0('Qualis: ', field.qualis))
print(p)
```

We can also use `dplyr` to do some simple assessment of academic productivity:

```{r}
library(dplyr)

my.tab <- tpublic.published %>%
  group_by(name) %>%
  summarise(n.papers = n(),
            max.SJR = max(SJR, na.rm = T),
            mean.SJR = mean(SJR, na.rm = T),
            n.A1.qualis = sum(qualis == 'A1', na.rm = T),
            n.A2.qualis = sum(qualis == 'A2', na.rm = T),
            median.authorship = median(as.numeric(order.aut), na.rm = T ))

knitr::kable(my.tab)
```
