---
title: 'New package: GetBCBData'
subtitle: "Importing datasets from BCB (Banco Central do Brasil)"
author: "Marcelo S. Perlin"
date: '2019-04-15'
draft: false
tags: ["R", "GetBCBData"]
categories: ["R", "GetBCBData"]
---

```{r, echo = FALSE}
knitr::opts_chunk$set(message = FALSE, cache = TRUE)
```

The Central Bank of Brazil (BCB) offers access to the SGS system (sistema gerenciador de series temporais) with a official API available [here](http://www.bcb.gov.br/?sgs).

Over time, I find myself using more and more of the available datasets in my regular research and studies. Last weekend I decided to write my own API package that would make my life (and others) a lot easier. 

Package GetBCBData can fetch data efficiently and rapidly: 

- Use of a caching system with package `memoise` to speed up repeated requests of data;
- Users can utilize all cores of the machine (parallel computing) when fetching a large batch of time series;
- Allows the choice for format output: long (row oriented, tidy data) or wide (column oriented)
- Error handling internally. Even if requested series does not exist, the function will still return all results.

## Installation

```{r, eval=FALSE}
# CRAN (official release) - IN CHECK
install.packages('GetBCBData')

# Github (dev version)
devtools::install_github('msperlin/GetBCBData')
```


## A simple example

Let's have a look at unemployment rates around the world. After searching for the ids in the [SGS system](http://www.bcb.gov.br/?sgs), we find the ids for 6 countries and set it as input `id`. 

Now, let's download the data with `GetBCBData`:

```{r, message=FALSE}
#devtools::install_github('msperlin/GetBCBData')

library(GetBCBData)
library(tidyverse)

my.countries <- c('Germany', 'Canada', 'USA', 
                  'France', 'Italy', 'Japan')

my.ids <- c(3785:3790)

names(my.ids) <- paste0('Unemp. rate - ', my.countries)

df.bcb <- gbcbd_get_series(id = my.ids ,
                           first.date = '2000-01-01',
                           last.date = Sys.Date())

glimpse(df.bcb)

p <- ggplot(df.bcb, aes(x = ref.date, y = value) ) +
  geom_line() + 
  labs(title = 'Unemploymnent Rates Around the World', 
       subtitle = paste0(min(df.bcb$ref.date), ' to ', max(df.bcb$ref.date)),
       x = '', y = 'Percentage*100') + facet_wrap(~series.name)


print(p)
```



