---
title: 'New package: GetQuandlData'
author: "Marcelo S. Perlin"
date: "2019-10-01"
subtitle: Quandl and tidyverse
draft: no
tags:
- R
- GetQuandlData
categories:
- R
- GetQuandlData
---

```{r, echo = FALSE}
knitr::opts_chunk$set(message = FALSE, cache = FALSE, eval= TRUE)
```

## Introduction

[Quandl](https://www.quandl.com/search) is one of the best platforms for finding and downloading financial and economic time series. The collection of free databases is solid and I use it intensively in my research and class material.

But, a couple of things from the native package `Quandl` always bothered me:

- Multiple data is always returned in the wide (column oriented) format (why??);
- No local caching of data;
- No control for importing error and status;
- Not easy to work within the `tidyverse` collection of packages

As you suspect, I decided to tackle the problem over the weekend. The result is package `GetQuandlData`. This is what it does differently:

- It uses the json api (and not the Quandl native function), so that some metadata is also returned;
- The resulting dataframe is always returned in the long format, even for multiple series;
- Users can set custom names for input series. This is very useful when using along `ggplot` or making tables;
- Uses package `memoise` to set a local caching system. This means that the second time you ask for a particular time series, it will grab it from your hard drive (and not the internet);
- Always compares the requested dates against dates available in the platform.


## Installation

```{r, eval=FALSE}
# not in CRAN yet (need to test it further)
#install.packages('GetQuandlData')

# from github
devtools::install_github('msperlin/GetQuandlData')
```

### Example 01 - Inflation in the US

Let's download and plot information about inflation in the US:

```{r}
library(GetQuandlData)
library(tidyverse)

my_id <- c('Inflation USA' = 'YALE/SP_CPI')
my_api <- readLines('~/GDrive/98-pass-and-bash/.quandl_api.txt') # you need your own API (get it at https://www.quandl.com/sign-up-modal?defaultModal=showSignUp>)
first_date <- '2005-01-01'
last_date <- Sys.Date()

df <- get_Quandl_series(id_in = my_id, 
                        api_key = my_api, 
                        first_date = first_date,
                        last_date = last_date, 
                        cache_folder = tempdir())

glimpse(df)
```

As you can see, the data is in the long format. Let's plot it:

```{r}
library(tidyverse)

p <- ggplot(df, aes(x = Year, y = value/100)) + 
  geom_col() + 
  labs(y = 'Inflation (%)', 
       x = '',
       title = 'Inflation in the US') + 
  scale_y_continuous(labels = scales::percent)

p
```

Beautiful!


### Example 02 - Inflation for many countries

Next, lets have a look into a more realistic case, where we need inflation data for several countries:

First, we need to see what are the available datasets from database `RATEINF`:

```{r}
library(GetQuandlData)
library(tidyverse)

db_id <- 'RATEINF'
my_api <- readLines('~/GDrive/98-pass-and-bash/.quandl_api.txt') # you need your own API

df <- get_database_info(db_id, my_api)

knitr::kable(df)
```

Nice. Now we only need to filter the series with YOY inflation:

```{r}
idx <- stringr::str_detect(df$name, 'Inflation YOY')
df_series <- df[idx, ]
```

and grab the data:

```{r}
my_id <- df_series$quandl_code
names(my_id) <- df_series$name
first_date <- '2010-01-01'
last_date <- Sys.Date()

df_inflation <- get_Quandl_series(id_in = my_id, 
                                  api_key = my_api,
                                  first_date = first_date,
                                  last_date = last_date)

glimpse(df_inflation)
```

And, finally, an elegant plot:

```{r}
p <- ggplot(df_inflation, aes(x = ref_date, y = value/100)) + 
  geom_col() + 
  labs(y = 'Inflation (%)', 
       x = '',
       title = 'Inflation in the World',
       subtitle = paste0(first_date, ' to ', last_date)) + 
  scale_y_continuous(labels = scales::percent) + 
  facet_wrap(~series_name)

p
```


