---
title: 'B3 is NOT shutting down its ftp site, for now..'
subtitle: "GetHFData still works!"
author: "Marcelo S. Perlin"
date: '2019-08-08'
draft: false
tags: ["R", "GetHFData"]  
categories: ["R", "GetHFData"]
---

```{r, echo = FALSE}
knitr::opts_chunk$set(message = FALSE, cache = TRUE, eval=FALSE)
```

**Update 2019-08-09: The shutdown is just postponed to 2019-11-14. See the official release [here](http://www.b3.com.br/pt_br/noticias/portais-8AE490C86BC98316016BDD1E7BDF14AB.htm?fbclid=IwAR1IY_ILkhkWwsBxFWH4pQEBx-OqxcEDxfmg22Ii5tep_dtzyKw9casag5o).**

Surprise, surprise. B3's ftp site is still up and running. 

Following previous [post](https://www.msperlin.com/blog/post/2019-07-01-ftp-shutdown/) regarding the shutdown of B3's ftp site and its impact over GetHFData, I'm happy to report that the site is up and running.  

We can check it with code:

```{r, message=FALSE, cache = TRUE}
library(GetHFData)
library(tidyverse)

df.ftp <- ghfd_get_ftp_contents(type.market = 'equity')

# check time difference
max(df.ftp$dates) - min(df.ftp$dates)
```

Let's download some trade data:

```{r, cache = TRUE}
df.trades <- ghfd_get_HF_data(my.assets = 'PETR3', 
                              type.market = 'equity', 
                              first.date = max(df.ftp$dates)-3, 
                              last.date = max(df.ftp$dates),
                              type.data = 'trades', type.output = 'agg', 
                              first.time =  '11:00:00', last.time = '18:00:00', dl.dir = tempdir()
                              )
```

And check it out: 

```{r}
glimpse(df.trades)
```

Its working fine. The amount of data available at the ftp is more than necessary for research and class material. 

I'm not really sure what happened. It could be a simple delay to the shutdown. Lets keep our fingers crossed. 


