---
title: "Book slides - Analyzing Financial and Economic Data with R"
author: "Marcelo S. Perlin"
date: '2020-02-25'
draft: false
subtitle: ' '
tags: ["R", "afedR"]
categories: ["R", "afedR"]
---


```{block, type='alert alert-warning'}
2021-02-28 -- Important: this blog post is deprecated. With the [last revision](https://www.msperlin.com/post/2021-02-28-afedr-revision-2021/) in march 2021, the Rmd slides are no longer available within the book material.
```

The slides for my newly released book [Analyzing Financial and Economic Data with R](https://www.msperlin.com/post/2020-01-15-afedr-ed2-announcement/) are finally ready! I apologize for keep you guys waiting. 

The slides are available as independent .Rmd files for all book chapters including: 

```{r, echo=FALSE}
my_dir <- '~/Dropbox/06-My Books/02-Analyzing Financial and Economic Data with R/Edition 2/03-Slides/'
my_f <- list.files(path = my_dir, pattern = '.Rmd')
my_f
```

All content is released with a generous MIT license, so fell free to use and edit the files as you wish.

You can download the slides and other book material with the following code:

```{r, eval=FALSE}
if (!require(devtools)) install.packages("devtools")

devtools::install_github('msperlin/afedR')

# ignore warning messages (long paths..)
afedR::afedR_get_book_files(path_to_copy = '~')
```

Alternatively, you can download the files directly from [github](https://github.com/msperlin/afedR/tree/master/inst/extdata).

The bundle also includes other teaching material that may help conduct a complete R tutorial:

- Solutions to end of chapter exercises

- All R Code used in the book

- Dynamic exercises (package exams)

- Data files

If you liked the material, please consider purchasing the book and **leaving your feedback at [Amazon](https://www.amazon.com/dp/B084LSNXMN)**. Your oppinion is very important for promoting the book and help others learn more about R and RStudio. As an author, I certainly appreciate the gesture and will take it as a motivating factor for future editions of the book. 
