---
title: "GetFREData available in CRAN!"
author: "Marcelo S. Perlin"
date: '2021-04-06'
tags:
- R
- GetFREData
categories:
- R
- GetFREData
draft: false
---

I'm happy to report that package `GetFREData` is now available in CRAN. This R package serves as an interface to **all** corporate datasets available in the FRE system, a vast and **official** repository of information about many different corporate events. All companies listed at B3 -- Brazilian stock exchange -- must report to FRE any significant change in their corporate structure. You can find more details about what is available in FRE in its [web interface](https://www.rad.cvm.gov.br/ENETCONSULTA/frmGerenciaPaginaFRE.aspx?NumeroSequencialDocumento=102618&CodigoTipoInstituicao=2).

The R package fetches data from the [CVM ftp](http://dados.cvm.gov.br/dados/CIA_ABERTA/DOC/FRE/), downloads and parses the xml files, and output several tables as a `list`. The corporate data includes (since 2010):

- List of stockholders
- All capital issues 
- Stock value over years
- Compensation of boards and directors
- Composition of boards and committees
- Family relations within the company
- List of companies related to family members
- Stock details
- Intangible details
- Auditing details
- Dividends details


Historical parsed data between 2010 and 2019 is available for download in my [personal site](https://www.msperlin.com/data/).


## Installation

```{r, eval=FALSE}
# CRAN (stable)
install.packages('GetFREData')

# github (development)
if (!require(devtools)) install.packages('devtools')
if (!require(GetFREData)) devtools::install_github('msperlin/GetFREData') 
```

## Example of usage

```{r}
library(GetFREData)
library(tidyverse)

search_company('grendene', 
               cache_folder = tempdir())

l_fre <- get_fre_data(companies_cvm_codes = 19615,
                      fre_to_read = 'last',
                      first_year = 2020,
                      last_year = 2020, 
                      cache_folder = tempdir())
                      
glimpse(l_fre)
```


