---
title: "Compiling Book Exercises to pdf | html | Moodle | Blackboard"
author: "Marcelo S. Perlin"
date: '2021-02-28'
tags:
- R
- afedR
categories:
- R
- afedR
draft: no
---

```{r, echo = FALSE}
knitr::opts_chunk$set(message = FALSE, eval=TRUE)

link_full_compiled_html <- 'https://www.msperlin.com/files/afedr-files/Introduction to R_Robert Engle_Ver 03.html'
link_book_web <- 'https://www.msperlin.com/adfeR/'
exams_link <- 'http://www.r-exams.org/'
```

The revised second edition of [Analylzing Financial and Economic Data with R](`r link_book_web`) presents more than 100 exercises at the end section of all chapters. All exercises are freely available in the [`exams` format](`r exams_link`), meaning that any R tutor can export the same exercises to pdf, html or _e-learning_ platforms. In this post I'll show how to compile exercises to pdf, html, _Moodle_ and _blackboard_.


# Installation 

The first step is to install package `afedR` with `devtools`:

```{r, eval=FALSE}
devtools::install_github('msperlin/afedR')
```

Another requirement is a working Latex instalation. For that, use `tinytex`:

```{r, eval=FALSE}
tinytex::install_tinytex()
```


# Compiling Exercises

## How it works?

All book exercises in the `exams` format: each exercise is a .Rmd file containing code, exercise text and solution. The files themselves can be found in the installation directory of the package:

```{r}
eoc_dir <- afedR::get_EOC_dir()

eoc_chapters <- fs::dir_ls(eoc_dir)
basename(eoc_chapters)
```

Each folder will have several exercises. Let's try one out:

```{r}
eoc_files <- fs::dir_ls(eoc_chapters[1])
basename(eoc_files)
```

We can also read one of the files to show the strucuture of the exercise in code and text:

```{r}
readLines(eoc_files[1])
```

Basically, we define all sections of a question -- text, solution, alternatives -- using a .Rmd template. Again, you can find more details about using package `exams` in its own [website](`r exams_link`).

Be aware that, all exams .rmd files available within `afedR` are self contained and you can export and compile them directly from `exams`. An easy way to copy all exercise files to your local folder is using function `afedR::path_to_copy`:

```{r, message=TRUE}
# copy to "documents" folder
afedR::copy_book_files(path_to_copy = '~')
```
All book files -- data, code and exercises -- are now available at your "Documents" folder (shorcut of `~`).

## Compiling to pdf

For pdf compilation, you'll need:

- name of students (will be printed in pdf)
- students ids (I usually use their university card number)
- Chapters to include
- Exercise name
- Course name

And use the following code

```{r, message=FALSE}
library(afedR)

names_students <- c('Michael Peterling', 'John Aspper', 'Mr. Beans')
ids_students <- 1:length(names_students) # probably id card?
chapters <- 1:3 # chapters from 1 to 13
dir_output <- file.path(tempdir(), 'pdf-example')


df_exams <- compile_pdf_exercises(students_names = names_students, 
                                  students_ids = ids_students, 
                                  chapters_to_include = chapters,
                                  dir_out = dir_output)
```

The output of `compile_pdf_exercises` is a table with the correct answers for `schoice` and `num` type of questions:

```{r}
glimpse(df_exams)
```


After compilation, all pdf files are available at folder `dir_output`:

```{r}
fs::dir_ls(dir_output)
```

The final result will be  as follows:

<blockquote class="imgur-embed-pub" lang="en" data-id="a/wVAd8Xr" data-context="false" ><a href="//imgur.com/a/wVAd8Xr"></a></blockquote><script async src="//s.imgur.com/min/embed.js" charset="utf-8"></script>


## Compiling to html

You can also compile to a html file using `afedR::compile_html_exercises`:

```{r, message=FALSE, eval=TRUE}
library(afedR)

names_students <- c('Michael Peterling', 'John Aspper', 'Mr. Beans')
ids_students <- 1:length(names_students) # probably id card?
chapters <- 1:3 # chapters from 1 to 13
dir_output <- file.path(tempdir(), 'html-example')

df_exams <- compile_html_exercises(students_names = names_students, 
                                   students_ids = ids_students, 
                                   chapters_to_include = chapters,
                                   dir_out = dir_output)
```

An example of full compiled html file is available [here](`r link_full_compiled_html`).


## Exporting to Moodle

You can also export to _e-learning_ platforms such as Moodle. The process is quite simple as `exams` package does the heavy work:

```{r}
require(afedR)
my_eoc_dir <- afedR::get_EOC_dir()

available_chapters <- fs::dir_ls(my_eoc_dir)
exercise_files <- fs::dir_ls(available_chapters[1])
dir_output <- file.path(tempdir(), 'moodle-test')

exams::exams2moodle(file = exercise_files, 
                    name = 'TestingMoodle', 
                    dir = dir_output)

fs::dir_ls(dir_output)
```

You can later import this .xml file in your Moodle class. 

## Exporting to Blackboard

```{r}
require(afedR)
my_eoc_dir <- afedR::get_EOC_dir()

available_chapters <- fs::dir_ls(my_eoc_dir)
exercise_files <- fs::dir_ls(available_chapters[1])
dir_output <- file.path(tempdir(), 'blackboard-test')

exams::exams2blackboard(file = exercise_files, 
                        name = 'TestingBlackBoard', 
                        dir = dir_output)

fs::dir_ls(dir_output)
```

This .zip file contains all exercises of chapter 01 and can be imported in your blackboard account. 
