---
title: "bookdown + exams + webex"
subtitle: Compiling reproducible exercises for your book
author: "Marcelo S. Perlin"
date: '2021-03-18'
tags:
- R
- afedR
- exams
- bookdown
- webex
categories:
- R
- afedR
- exams
- bookdown
- webex
draft: no
---

It's been three years since I've been using package `bookdown` for compiling and distributing [three different books](https://www.msperlin.com/blog/publication/#5) in Amazon and the [web](https://www.msperlin.com/afedR/). It helped me greatly in all my book projects and I'm always **grateful** to Yihui Xie for providing such a useful tool at the [right time](https://www.msperlin.com/blog/post/2017-02-16-writing-a-book/). 

However, `bookdown` offers no support for chapter exercises of any sort. While you can write exercises in plain RMarkdown, it is not a good solution for a long term project such as a technical book. When writing the latest edition of [Analyzing Financial and Economical Data with R](https://www.msperlin.com/afedR/), I aimed for a  work cycle where the 100 plus exercises and their solutions were reproducible and easier to maintain. 

Meanwhile, package `exams` provides a [framework](http://www.r-exams.org/) to produce exercises in a reproducible setup, making it possible to export the exercises to any given format such as pdf or html, or even _e-learning_ platforms such as Moodle and Blackboard. I use `exams` extensively in all my university classes and it works like a charm! 

So, while writing afedR, I worked towards finding a way to bring the two technologies closer to each other, which is what I'll report in this blog post. Here are the main advantages of this setup:

- The content of book exercises, their solution and explanation in a single location (no more fidling with different folders).
- Dynamic output for html, with buttons and solutions available at a single click. 
- Exportable exercises for classes (see this [blog post](https://www.msperlin.com/blog/post/2021-02-28-dynamic-exercises-afedr/)). You can export the same exercises to pdf or Moodle, for example. 


## `bookdown` + `exams` + `webex` 

First and foremost, the main part of the hack is to realize that any exercises in a .Rmd file [can be broken](https://stackoverflow.com/questions/62315622/using-r-exams-in-bookdown-document-especially-for-html-output/66012008#66012008) into a list using `exams::xexams`. Let's use an example from the book, with the first three exercises of chapter 01:

```{r}
# example from book 
afedR::copy_book_files(path_to_copy = tempdir())

# temp folder with exercises
eoc_dir <- file.path(tempdir(), 'afedR files/eoc-exercises/')

# select exercises
my_exercises <- list.files(eoc_dir, pattern = '*.Rmd', full.names = TRUE)
my_exercises <- my_exercises[1:3]

# break it down
my_l <- exams::xexams(my_exercises)

# check it
dplyr::glimpse(my_l)
```

As an example, in this list you can see the main text of the question 01 in slot `l_out$exam1$exercise1$question`:

```{r}
my_l$exam1$exercise1$question
```

And the solution at `my_l$exam1$exercise1$solution`

```{r}
my_l$exam1$exercise1$solution
```

In my case, I wanted the html version of the book to have all the solutions hidden by a clickable button -- just like in `webex` --  while the pdf and ebook would only have the text of the questions. Here are the functions I used:

```{r}
compile_eoc_exercises <- function(files_in, type_doc) {
  my_counter <<- 1

  if (is.null(type_doc)) {
    type_doc = 'html'
    #type_doc = 'latex'
  }

  for (i_ex in files_in) {
    exercise_to_html(i_ex, my_counter = my_counter,
                     type_doc)

    my_counter <<- my_counter +1
  }

  return(invisible(TRUE))
}

exercise_to_html <- function(f_in, my_counter, type_doc) {

  require(exams)
  require(webex)
  require(tidyverse)

  text_pre_solution <- paste0('To reach the same result, you must execute the code below. ',
                           'For that, open a new R script in RStudio (Control+shift+N), ',
                           'copy and paste the code, and execute it whole by pressing ',
                           'Control+Shift+Enter or line by line with shortcut ',
                           'Control+Enter.')

  my_dir <- file.path(tempdir(), basename(tempfile()))
  dir.create(my_dir)

  suppressMessages({
  l_out <- exams::xexams(f_in, driver = list(sweave = list(png = TRUE)),
                         dir = my_dir)

  })

  exercise_text <- paste0(l_out$exam1$exercise1$question, collapse = '\n')
  alternatives <- l_out$exam1$exercise1$questionlist
  correct <- l_out$exam1$exercise1$metainfo$solution
  solution <- paste0(l_out$exam1$exercise1$solution,
                     collapse = '\n')
  ex_type <- l_out$exam1$exercise1$metainfo$type

  if (type_doc %in% c('latex', 'epub3')) {

    my_str <- str_glue('\n\n {sprintf("%02d", my_counter)} - {exercise_text} \n\n ')

    if (ex_type == 'schoice') {
      n_alternatives <- length(alternatives)

      for (i_alt in seq(1, n_alternatives)) {
        my_str <- paste0(my_str,
                         letters[i_alt],') ', alternatives[i_alt],
                         '\n')
      }

    }

    cat(my_str)

    return(invisible(TRUE))

  } else if (type_doc == 'html') {

    if (ex_type == 'schoice') {
      vec_mcq <- sample(
        c(alternatives[!correct],
          answer = alternatives[correct])
      )

      my_answers_text <- str_glue('<br> Solution: {mcq(vec_mcq)}')
      numeric_sol <- alternatives[correct]
      text_sol <- str_glue('The solution is {numeric_sol}. {text_pre_solution}')

    } else if (ex_type == 'num') {

      numeric_sol <- correct
      my_answers_text <- str_glue('<br><br> Your Answer: {fitb(correct)}')
      text_sol <- str_glue('The solution is {numeric_sol}. {text_pre_solution}')

    } else if (ex_type == 'string') {
      my_answers_text <- ''
      numeric_sol <- ''

      if (stringr::str_detect(solution,
                              '```text')) {
        text_sol <- paste0('In order to execute the code, open a new R script in RStudio (Control+shift+N), ',
                           'copy and paste the code, and execute it whole by pressing ',
                           'Control+Shift+Enter or line by line with shortcut ',
                           'Control+Enter.')

      } else {
        text_sol <- ''
      }

    }


    my_str <- paste0('\n\n <hr> \n',
                     webex::total_correct(), '\n',
                     '### Q.', my_counter, '{-} \n',
                     exercise_text, '\n',
                       my_answers_text)


    temp_id <- basename(tempfile(pattern = 'collapse_'))
    sol_str <- str_glue(
      ' <div style="text-align: left; margin-top: 2px; padding: 13px 0 10px 0;"><p><button class="btn btn-primary" type="button" data-toggle="collapse" data-target="#{temp_id}" aria-expanded="false" aria-controls="collapseExample">
    Solution
  </button> </p> </div>

<div class="collapse" id="{temp_id}">
{text_sol}
  <div class="card card-body">
    {solution}
  </div>
</div>')

    cat(paste0(my_str, '\n' ,
               sol_str))

  }

  return(invisible(TRUE))

}
```


### Html Exercises 

The html output for the selected three exercises is given next. Do notice that the correct solution is **not highlighted** in this blog post due to the lack of css and javascript. In the [final result](https://www.msperlin.com/afedR/introduction.html#exercises) you'll see that it works correctly. Also, you'll need to set `results='asis'` in the knitr options of the chunk (the code output pure html).

```{r, results='asis', message=FALSE}
compile_eoc_exercises(my_exercises, type_doc = 'html')
```


### Pdf/Ebook Exercises 

And for latex (pdf) and epub3 (ebook), the result is:

```{r, results='asis', message=FALSE}
compile_eoc_exercises(my_exercises, type_doc = 'latex')
```

As you can see, it works great. So, at the end of each chapter I simply called function `compile_eoc_exercises()` with the knit chunk options `results='asis'` and `echo=FALSE`. Moreover, object `my_engine` is set as `my_engine <- knitr:::pandoc_to()`, which will figure out the format within the compilation of the book:

<blockquote class="imgur-embed-pub" lang="en" data-id="a/lCDuvZg" data-context="false" ><a href="//imgur.com/a/lCDuvZg"></a></blockquote><script async src="//s.imgur.com/min/embed.js" charset="utf-8"></script>


## Conclusion

Its is amazing how much we can accomplish by learning and mixing different technologies. In this case, I used R, Latex, html, javascript and css to bundle reproducible and dynamic exercises for my book. You can find examples of the final output in [html](https://www.msperlin.com/afedR/introduction.html#exercises), [pdf](https://www.msperlin.com/afedR/afedR_ed02-ONLINE.pdf) and [ebook](https://www.msperlin.com/afedR/afedR_ed02-ONLINE.epub). 

If you're trying it for you own book, make sure to add the correct .js and .css files to the html compilation. In my case, I used [my_javascript.js](https://www.msperlin.com/blog/post/my_javascript.js) and [style_html.css](https://www.msperlin.com/blog/post/style_html.css). 

