---
title: "Compiling book exercises"
author: "Marcelo S. Perlin"
date: "2023-03-10"
subtitle: Exporting exercises to pdf | Moodle | Blackboard
tags:
- R
- afedR
categories:
- R
- afedR
draft: yes
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(message = FALSE, eval=TRUE)
options(knitr.duplicate.label = "allow")

exerc_folders <- purrr::map_chr(
  afedR3::exercises_dir_list(),
 afedR3::exercises_dir_get
 )

n_exerc <- length(fs::dir_ls(exerc_folders, recurse = TRUE))

my_links <- afedR3::links_get()
link_book_web <- my_links$book_amazon_ebook
exams_link <- 'http://www.r-exams.org/'
```

The third edition of [Analyzing Financial and Economic Data with R](`r link_book_web`) provides a total of `r n_exerc` end-of-chapter exercises. All activities are freely available in the [`exams` format](`r exams_link`), meaning that any R tutor can export the same exercises and solutions to use in their own class. In this post I'll show how to compile exercises to pdf, html, _Moodle_ and _blackboard_.


# Installation 

The first step is to install package `afedR3` with `devtools`:

```{r req, eval=FALSE}
if (!require(devtools)) install.packages('devtools')

devtools::install_github('msperlin/afedR3')
```

Another requirement is a working Latex instalation. For that, use `tinytex`:

```{r tinytex, eval=FALSE}
tinytex::install_tinytex()
```


# Compiling Exercises

## How it works?

All book exercises are written in the `exams` format: each exercise is a .Rmd file containing code, exercise text and solution. The files themselves can be found in the installation directory of the book package, and each folder contains the exercise for a particular chapter:

```{r eoc}
eoc_dir <- afedR3::get_EOC_dir()

eoc_chapters <- fs::dir_ls(eoc_dir)
basename(eoc_chapters)
```

Let's have a deeper look at the exercises of the first chapter:

```{r eoc2}
eoc_files <- fs::dir_ls(eoc_chapters[1])
basename(eoc_files)
```

We can also read one of the files to show the structure of the exercise in code and text:

```{r read-eoc}
readLines(eoc_files[1])
```

In a nutshell, we define all sections of a question -- text, solution, alternatives -- using a .Rmd template. Again, you can find more details about using package `r afedR3::format_pkg_text("exams")` in its [website](`r exams_link`).


## Compiling to pdf

To help tutors compiling their own exercises, I wrote function `r afedR3::format_fct_ref("afedR3", "compile_pdf_exercises")`. You'll need the following information to use it:

- (REQUIRED) name of students (will be printed in pdf)
- (OPTIONAL) students ids (I usually use their university card number)
- (OPTIONAL) Chapters to include (e.g 1:3)
- (OPTIONAL) Exercise name (e.g. Activity I, Exercise II, ..)
- (OPTIONAL) Course name (e.g. Tutorial in R)

Here's an example:

```{r compiling-pdf}
library(afedR3)

names_students <- c('Michael Peterling', 'John Aspper', 'Mr. Beans')
ids_students <- sample(1:1000, length(names_students)) # probably id card?
class_name <- "Introduction to R"
exercise_name <- "Activity 01"
chapters <- 1:3 # chapters from 1 to 13
dir_output <- fs::file_temp('pdf-example_')

df_exams <- compile_pdf_exercises(students_names = names_students, 
                                  students_ids = ids_students, 
                                  class_name = class_name,
                                  exercise_name = exercise_name,
                                  chapters_to_include = chapters,
                                  dir_out = dir_output)
```

The output of `compile_pdf_exercises` is a table with the correct answers for `schoice` and `num` type of questions:

```{r solutions}
glimpse(df_exams)
```

After compilation, all pdf files are available at folder `dir_output`:

```{r pdf-files}
fs::dir_ls(dir_output)
```

The final result will be  as follows:

<blockquote class="imgur-embed-pub" lang="en" data-id="a/LO9xafp" data-context="false" ><a href="//imgur.com/a/LO9xafp"></a></blockquote><script async src="//s.imgur.com/min/embed.js" charset="utf-8"></script>


## Exporting to Moodle

You can also export to _e-learning_ platforms such as Moodle. The process is quite simple as `exams` package does all the heavy work:

```{r moodle}
library(afedR3)

eoc_folders <- afedR3::get_EOC_dir()

available_chapters <- afedR3::exercises_dir_list()

exercises_folders <- purrr::map_chr(
  available_chapters[1:3],
  afedR3::exercises_dir_get
)

exercises_files <- fs::dir_ls(exercises_folders)

dir_output <- fs::file_temp('moodle-test_')

exams::exams2moodle(file = exercises_files, 
                    name = 'MOODLE_afedR-eoc-chapters_01-03', 
                    dir = dir_output)

fs::dir_ls(dir_output)
```

The resulting .xml file can be imported in the database of any Moodle class you have access. 

## Exporting to Blackboard

Likewise, exporting to [Blackboard](https://www.blackboard.com/) is simple:

```{r blackboard}
library(afedR3)
library(tth) # required for bb 

dir_output <- dir_output <- fs::file_temp('blackboard-test_')

exams::exams2blackboard(file = exercises_files, 
                        name = 'BB_afedR-eoc-chapters_01-03', 
                        dir = dir_output)

fs::dir_ls(dir_output)
```

This .zip file contains all exercises of chapters 01, 02 and 3, and can be imported in your blackboard account. 
